Stable Hot Spot Analysis (Draft)

### 2017-06-13 paper-main.tex
first draft related work TODO: Citations ...

\begin{abstract}
Hotspot analysis is essential for geo-statistics.
It supports decision making by detecting points
as well as areas of interest in comparison to their
neighbourhood. However, these methods are dependent
on different parameters, ranging from the resolution
of the study area to the size of their neighbourhood.
This dependence can lead to instabilities of the detected
hotspots, where the results can highly vary between
different parameters. A decision maker can therefore ask
how valid the analysis actually is.
%
In this study, we examine the impact of key parameters
on the stability of the hotspots, namely the size of
the neighbourhood, the resolution and the size of the
study area, as well as the influence of the ratio between
those parameters.
%
We compute the hotspots with the well known Getis-Ord ($G^*$)
statistic as well as its modification, the \emph{Focal $G^*$} statistic.
We measure the stability of the hotspot analysis using
a recently introduced \emph{stability of hotspots} metric (SoH)
and compare the results to intuitive visual analysis.
%
We evaluate the results on real world data with the well-known
yellow cab taxi data set from New York, Manhattan.
Our results indicate a negative impact on the stability with
an increase of the size of the neighbourhood as well as
a reduction of the size of the study area, regardless of
the resolution.
\end{abstract}

\section{Introduction}
The goal of hotspot analysis is the detection and identification of interesting areas. It achieves this goal by computing statistically significant deviations from the mean value of a given study area. This allows a decision maker to easily identify those areas of interest and allows further focus in sub-sequential data analysis or the decision focus. Typical applications range from crime detection over identification of disease outbreaks to urban heat islands. In such applications, scarce resources are then often applied in only those identified hot spots or used as the basis for the allocation. The general approach is an unsupervised learning method similar to a cluster analysis.

@@

\begin{itemize}
\item goal Here: some cluster analysis
\item typical clustering apporaches -> phd symposium as examples (possible to cite =) )
\end{itemize}

The goal of an analysis of temperatures in a city is to find the most
interesting, significant areas: Hot spots~\cite{Martin.2015}. This goal is
similar to the hot spot analysis in the field of geo-statistics. One of the most
fundamental approach is Moran´s I~\cite{MoranI}. There it is tested whether or
not a spatial dependency exists. This gives the information on global
dependencies in a data set. Upon this hypothesis test several geo-statistical
tests are based. The most well known are the Getis-Ord statistic~\cite{Ord.1995}
and LISA~\cite{Anselin.1995}. In both cases the general, the global statistic of
Moran´s I is applied in a local context. The goal is to detect not only global
values, but instead to focus on local hot spots and to measure the significance
of those local areas.

The local Getis-Ord statistic~\cite{Ord.1995} is defined as follow:
\begin{definition}[Getis-Ord $G^*_i$ statistic] \label{def:Gstar}
  Assuming a study area with $n$ measurements, let $X = [x_1, \ldots,
  x_n]$ be all values measured in this area. Let $w_{i,j}$ be a spatial weight
  between two points $i$ and $j$ for all $i,j \in \{ 1, \ldots, n\}$. The 
  Getis-Ord $G^*_i$ statistic is given as:
  \begin{equation} \label{eq:Gstar}
  G^*_i = \frac{
    \sum_{j=1}^{n}w_{i,j}x_j - \bar{X}\sum_{j=1}^{n}w_{i,j}
  }{
  S \sqrt{
    \frac{
      n \sum_{j=1}^{n}w_{i,j}^2 - (\sum_{j=1}^{n}w_{i,j})^2
    }{n-1}
  }
}
\end{equation}
where:
\begin{itemize}
  \item $\bar{X}$ is the mean of all measurements,
  \item $S$ is the standard deviation of all measurements.
\end{itemize}
\end{definition}

As it is known, this statistic creates a z-score, which denotes the significance
of an area in relation to its surrounding areas.

LISA is quite similar, as it is the local statistic for Moran's
I~\cite{Anselin.1995}, but the z-score has a different meaning. Apart from
$G^*_i$, LISA does not distinguish between cold spots and hot spots as it
assigns high z-score to most similar areas.

Two other well known method are the kernel density estimation~\cite{KDE_Hotspot}
and kriging~\cite{Kriging_Hotspot}. These do not provide significance levels.
Instead, they estimate the values for each location based on the rest of the
study area and a threshold value \cite{Thakali2015}. Therefore, results for
different areas are not comparable, especially in the case of differing
temperature distributions. Kriging was developed for the estimation of ore
deposits~\cite{krige1951statistical}, but today, applications for geo-temporal
forecasts with this approach can be found, e.g. for the city of
Zurich.~\footnote{https://r-video-tutorial.blogspot.de/2015/08/spatio-temporal-kriging-in-r.html}

All of the aforementioned methods use weights between pairs of points, usually
based on their geographical distance. However, in real applications, the points
are aggregated into rasters and the weights are represented as a weight matrix.
This allows for expressing the algorithms in terms of map algebra operations, a
term first coined by Dana~Tomlin~\cite{Tomlin1990} and computed in a distributed
fashion (e.g. using Geotrellis framework running on
Apache~Spark~\cite{github:Geotrellis}).

In this work, we focus on the Getis-Ord statistic applied to rasters of land
surface temperature. This enables us to transform the formula for the $G^*$
statistic into a computationally more efficient form. We then propose a
modification of the standard $G^*$ statistics, to increase the stability of the
hot spots found.

@@

For a data analyst, when exploring the data interactively by choosing different
filter sizes (weight matrices) or point aggregation strategies (pixel sizes), it
is important that the position and size of a hotspot changes in a predictable
manner. We formalize the intuition in our stability metric.

>>>>

\begin{abstract}
Hot spot analysis is essential for geo-statistics.
It supports decision making by detecting points
as well as areas of interest in comparison to their
neighbourhood. However, these methods are dependent
on different parameters, ranging from the resolution
of the study area to the size of their neighbourhood.
This dependence can lead to instabilities of the detected
hotspots, where the results can highly vary between
different parameters. A decision maker can therefore ask
how valid the analysis actually is.
%
In this study, we examine the impact of key parameters
on the stability of the hotspots, namely the size of
the neighbourhood, the resolution and the size of the
study area, as well as the influence of the ratio between
those parameters.
%
We compute the hotspots with the well known Getis-Ord ($G^*$)
statistic as well as its modification, the \emph{Focal $G^*$} statistic.
We measure the stability of the hotspot analysis using
a recently introduced \emph{stability of hotspots} metric (SoH)
and compare the results to intuitive visual analysis.
%
We evaluate the results on real world data with the well-known
yellow cab taxi data set from New York, Manhattan.
Our results indicate a negative impact on the stability with
an increase of the size of the neighbourhood as well as
a reduction of the size of the study area, regardless of
the resolution.
\end{abstract}

\section{Introduction}
The goal of hotspot analysis is the detection and identification of interesting areas. It achieves this goal by computing statistically significant deviations from the mean value of a given study area. This allows a decision maker to easily identify those areas of interest and allows further focus in sub-sequential data analysis or the decision focus. Typical applications range from crime detection over identification of disease outbreaks to urban heat islands. In such applications, scarce resources are then often applied in only those identified hotspots or used as the basis for the allocation. The general approach is an unsupervised learning method similar to a cluster analysis.

@@

\subsection{Quality of Clustering}
The problem of assessing the quality in unsupervised learning is well known. In the case of the k-mean algorithm, the quality of the clustering is mostly dependent on the value of the \emph{k} and a miss-specification can lead to highly irregular clusters. In a simple 2D clustering, they can be easily recognized by visual analysis, but in higher dimensionality, this is impossible. One method, to measure the quality of such a clustering is the compactness of the clusters, see e.g. Song 2010 (http://ieeexplore.ieee.org/document/5558958/?part=1). This enables the comparison between different clusters. Another possibility is the Silhouette Coefficient by Kaufman et Rousseeuw 1990. This metric measures the similarity of objects in a cluster in comparison to other clusters. For density based clustering, e.g. for DBSCAN, OPTICS (Ankerst et al 1999) gives a simple method to tune the essential parameter for this clustering. This is only a small overview of methods to influence and measure the quality of different clustering methods. But it shows that this problem is not easily solved and dependent on the chosen algorithm. To our knowledge, there does not exist a method to overall measure the stability of a clustering.

\subsection{Hot Spot Analysis}
The goal of hotspot analysis is the detection of interesting areas as well as patterns in spatial information. One of the most fundamental approach is Moran´s I~\cite{MoranI}. There it is tested whether or
not a spatial dependency exists. This gives the information on global
dependencies in a data set. Upon this hypothesis test several geo-statistical
tests are based. The most well known are the Getis-Ord statistic~\cite{Ord.1995}
and LISA~\cite{Anselin.1995}. In both cases the general, the global statistic of
Moran´s I is applied in a local context. The goal is to detect not only global
values, but instead to focus on local hotspots and to measure the significance
of those local areas. A more in depth overview of methods to identify and visualize spatial patterns and areas of interest can be found in Shekhar et al (2011).

@@

For a data analyst, when exploring the data interactively by choosing different
filter sizes (weight matrices) or point aggregation strategies (pixel sizes), it
is important that the position and size of a hot spot changes in a predictable
manner. We formalize the intuition in our stability metric.
